this script is to recode the parameters in Brown et al. (2018) for WEIRDO to enable the automatic fitting of xf and cf

```{r setup, include=FALSE}
rm(list = ls())
################
# define functions used 
######################

calc_HC <- function(cf, xf, rr, np){
  #function to calculate hydraulic conductivity from cf xf
  
  # create a dataframe 
  df <- cbind.data.frame(rep(cf, length(rr)),rr,rep(xf, length(rr)), np)
  volumetricFlow <- apply(df, 1, function(x){
    aa <-as.numeric(x)
    # vp <- cf* r^xf * np
    y <- aa[1]*(aa[2]^aa[3])*aa[4]
    y
  })
  poreCapillarity <- volumetricFlow/1e6*3600
  kthetas <- cumsum(rev(poreCapillarity))
  out <- data.frame('meanPoreRadius' = rr, 
                    'No_of_Pores' = np,
                    'PoreCapillarity' = poreCapillarity,
                    'HydraulicConductivity' = rev(kthetas))
  return(out)
}

optimCFXF <- function(cfxf, rr, np, ksat){
  # function to be optimised to obtain best xf and cf values
  cf <- cfxf[1]
  xf <- cfxf[2]
  # create a dataframe 
  df <- cbind.data.frame(rep(cf, length(rr)),rr,rep(xf, length(rr)), np)
  volumetricFlow <- apply(df, 1, function(x){
    aa <-as.numeric(x)
    # vp <- cf* r^xf * np
    y <- aa[1]*(aa[2]^aa[3])*aa[4]
    y
  })
  poreCapillarity <- volumetricFlow/1e6*3600
  kcalc_sat <- sum(poreCapillarity)
  out <- abs(ksat-kcalc_sat)
  return(out)
}

############################
# 1. model structure
# Pore characteristics, each profile layer contains 11 pore cohorts where similar sized pores are grouped together, with indices 0-9
# Dw is set when model processes
# model requires water contents at saturation, air-entry water potential head, drained upper limit, lower limit, and air dry
# theta at matric potential 0, variable (between -10 ~ -1000), -1000, -150000, -6.0E7 mm

# there are 10 pore cohorts per layer
# diameters of the boundaries for the 10 pore cohorts
poreBoundaries <- poreDiameterUpper <- c(3000, 1194, 475, 189, 75, 30, 8.6, 2.47, 0.707, 0.202, 0.0005)
poreDiameterLower <- c(poreDiameterUpper[2:11], 0)


# matric potential corresponding to SAT, DUL, LL, oven dry (mm)
# if no hbub data, assume it from the high sand soil in Brown 2018
# thetaS saturation was assumed the same as hbub

mPotential <- -c(1,10, 1000, 150000, 6.09E7)

```

```{r input, include=FALSE}
###general set up ------------
# get the soil attributes 
# (This is based on SoilSet 4592 for KParamFit (https://github.com/HamishBrownPFR/WEIRDO/blob/master/KParamFit.ipynb), no data for theta at AirDry, using 0.001 to ensure positive waterRetentionCurve)
#'SAT','wchBub','DUL', 'LL15',  'AirDry'
thetas <- c(0.45, 0.45, 0.38, 0.20, 0.001)

#setting psibub here
psibub <- -100


```

Fit a 5-point Hermite spline to derive the water retention curve for the soil profile, and use an optimisation routine to calculate these two empirical parameters based on the hydraulic conductivity at saturation, and the number of pores in each pore cohort

```{r opt, include=FALSE}
mPotential[2] <- psibub

# get water potential at pore boundaries
psiBoundary <- -30000/poreBoundaries

# get the mean pore radii of pore cohort
rr <- (poreDiameterUpper+poreDiameterLower)/4

####################
# fit hermite splines to soil-water characteristic and the conductivity curve
######################
# fit the 5-point hermite spline 
waterRetentionC <- splinefun(x = mPotential, y = thetas, method = 'monoH.FC')

# get thetas at pore boundaries using the fitted spline
thetas.upperBound <- waterRetentionC(psiBoundary)
thetas.lowerBound <- c(thetas.upperBound[2:11],0)
# get the water volume in each pore cohort
waterVolumePore <- thetas.upperBound-thetas.lowerBound

# calculate number of pores in each pore cohort
numbersPore <- waterVolumePore/((pi * rr^2)/10^12)

# The hydraulic-flow model (Aryal et al. 1999 implemented in Brown et al. 2018)
# to derive Cf and Xf

cfxf <- optim(c(5e-12, 3.5), optimCFXF, rr = rr, np = numbersPore, ksat = 0.2359)
while (cfxf$value != 0){
  cfxf <- optim(cfxf$par, optimCFXF, rr = rr, np = numbersPore, ksat = 0.2359)
}
cfxf <- cfxf$par

cfxf
```

Cf = `r cfxf[1]` Xf = `r cfxf[2]`

```{r spline, echo=FALSE}
# plot the spline to inspect
plot(log10(-mPotential), thetas, xlab = 'log10(-Suction)', 
     ylab = expression(theta*("mm"^3/'mm'^3)))
curve(waterRetentionC(-10^x, deriv = 0), add = T)
```

```{r k, echo=FALSE}

# plot hydraulic conductivity
df.hc <- calc_HC(cfxf[1], cfxf[2], rr, numbersPore)
plot(log10(-psiBoundary), df.hc$HydraulicConductivity, xlab = 'log10(-Suction)', 
     ylab = 'Hydraulic conductivity (mm/h)')
lines(log10(-psiBoundary), df.hc$HydraulicConductivity)

```
